```{r}
library(fpp3)
```

# 2.1 The tsibble object

Got to set up the tsibble (a time series enabled tibble) and specify which column holds the temporal bit with the `index` option.  

```{r}
y <- tsibble(
  Year = 2015:2019,
  Observation = c(123,39,78,52,110),
  index = Year
)
```

If not just year, need to do a little maintenance on the index


```{r}
z <- tibble(Month = c("2019 Jan", "2019 Feb", "2019 Mar","2019 Apr","2019 May"),
            Observation = c(50,23,34,30,25))

z %>% 
  mutate(Month = yearmonth(Month)) %>%
  as_tsibble(index = Month)
```

Options for mutating the index are:

- Annual - just need start:end
- Quarter - `tsibble::yearquarter()`
- Monthly - `tsibble::yearmonth()`
- Weekly - `tsibble::yearweek()`
- Daily - `lubridate::ymd()`
- HMS - `lubridate::ymd_hms()`

Multiple time series can be stored in a single tsibble

An example:

```{r}
olympic_running
```

And `tsibble` objects are perfectly happy working with dplyr, so munge away!

```{r}
PBS %>%
  filter(ATC2 == "A10") %>%
  select(Month, Concession, Type, Cost) %>%
  summarise(TotalC = sum(Cost)) %>%
  mutate(Cost = TotalC / 1e6) -> a10
```

Keys are another concept to be aware of with `tsibble`.  Essentially breaks the time series into other individual time series.

```{r}
prison <- readr::read_csv("https://OTexts.com/fpp3/extrafiles/prison_population.csv")

prison <- prison %>%
  mutate(Quarter = yearquarter(Date)) %>%
  select(-Date) %>%
  as_tsibble(key = c(State, Gender, Legal, Indigenous),
             index = Quarter)
```

Lastly, is the seasonal period.  The repeat interval for a given "season" or the number of observations in a year (e.g. monthly has 12, then it repeats).  Can have more that one seasonal pattern for more than weekly observations. Can specify more complex ones with `lubridate::period`.

# 2.2. Time plots

First ts plot!

```{r}
melsyd_economy <- ansett %>%
  filter(Airports == "MEL-SYD", Class == "Economy") %>%
  mutate(Passengers = Passengers/1000)
x <- autoplot(melsyd_economy, Passengers) +
  labs(title = "Ansett airlines economy class",
       subtitle = "Melbourne-Sydney",
       y = "Passengers ('000)")
x
```

Another plot

```{r}
autoplot(a10, Cost) +
  labs(y = "$ (millions)",
       title = "Australian antidiabetic drug sales")
```

# 2.4 Seasonal Plots

```{r}
a10 %>%
  gg_season(Cost, labels = "both") +
  labs(y = "$ (millions)",
       title = "Seasonal plot: Antidiabetic drug sales")
```

```{r}
vic_elec %>% 
  filter(year(Time) == 2012) %>%
  gg_season(Demand, period = "year") +
  theme(legend.position = "none") +
  labs(y="MWh", title="Electricity demand: Victoria")
```